# Reproducible Research: Peer Assessment 1
========================================
  by Pedro Rebelo  
github repo with RMarkdown source code:
 https://github.com/trashmanp1/RepData_PeerAssessment1  
SHA-1 hash


## Loading and preprocessing the data
=====================================
Loading the data is easy, just set the work directory to the directory where data is, unzip the data and load it

```{r}
setwd("/Users/pedrorebelo/RepData_PeerAssessment1")#path to directory where data is 
unzip("activity.zip")
dados <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Because in a day there are ``r (24*60)/5`` periods of 5 minutes, we must agregate the result by day, before take the mean and median:
```{r} 
passos_por_dia <- aggregate(dados[,"steps"],by=as.list(dados["date"]),FUN=sum,na.rm=TRUE)
names(passos_por_dia) <- c("date","steps")
media<-round(mean(passos_por_dia$steps)) #integer number of steps 
mediana<-median(passos_por_dia$steps)
```
 The general distribution is given by the hystogram:
```{r} 
hist(passos_por_dia$steps,breaks=round(log2(dim(passos_por_dia)[1])+1), #Sturges rule
      xlab="number of steps",
      main="Total number of steps per day") 
```

so, the mean is ``r media``steps per day, and the median is ``r mediana`` steps per day.

## What is the average daily activity pattern?
We can get the number of steps for the 5 minute intervals averaged over all  days

```{r} 
passos_por_5_min <- aggregate(dados[,"steps"],by = as.list(dados["interval"]),
                              FUN = mean, na.rm=TRUE)
plot(passos_por_5_min, type = "l",
      xlab="time (min)",
      ylab="mean steps",
      main="Daily 5 min average steps") 
```
 
 The 5-minute interval, on average across all the days in the dataset that contains the 

```{r} 
maximo<-passos_por_5_min$interval[which.max(passos_por_5_min$x)]
```

maximum number of steps begin at ``r maximo`` min.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
